# Missing values


```{r setup, include=FALSE}
# this prevents package loading message from appearing in the rendered version of your problem set
knitr::opts_chunk$set(warning = FALSE, message = FALSE,
                      echo = TRUE)
```

Our data consists of macroeconomic and financial information that don't contain missing values, mainly due to the structured nature of such information. To exemplify this, here we implement `visualize_missings` on the inflation data for Mexico from 2019 to 2021, and the US from 2012 to 2021 (monthly frequency).

```{r}
library(dplyr)
library(tidyr)
source("missing_visualization.R")

inflation_mx <- read.csv("Data/Inflation_MX_2019_2021.csv", stringsAsFactors = F)
inflation_us <- read.csv("Data/Inflation_US_2012_2021.csv", stringsAsFactors = F)

visualize_missings(inflation_mx)
visualize_missings(inflation_us)

```

Thus, we then proceed to showcase `visualize_missings` on the example dataset:

````{r fig.width=15, fig.height=10, fig.align="center"}

visualize_missings(fivethirtyeight::steak_survey,T)

visualize_missings(fivethirtyeight::steak_survey,F)


```
